Can Auxiliary Indicators Improve COVID-19 Forecasting and Hotspot Prediction?

### 2021-10-29 forecast/code/figures/compare-to-hub-ryan.Rmd
First push done
- Rewrite sig statement
- Rewrite abstract
- Pass over and rewrite muhc of intro
- Generally trying to describe the sience as much as possible,
  while maintaining language that I believe is accurate and not
  over-stating our conclusions

# Load our models ---------------------------------------------------------
```{r}
library(tidyverse)
ours <- readRDS(here::here("data", "results_honest_states.RDS")) %>%
  filter(forecaster != "AR3GSSAA3_Subset")
pop <- covidcast::state_census %>% select(ABBR, POPESTIMATE2019) %>%
  mutate(geo_value = tolower(ABBR)) %>% select(-ABBR)
# scale from prop to num
ours2 <- left_join(ours, pop, by = "geo_value") %>%
  mutate(wis = wis * POPESTIMATE2019, #/ 1e5 * 7,
         forecaster = recode(forecaster,
                             AR3 = "AR",
                             AR3CHCLI3 = "CHNG-CLI",
                             AR3CHCOV3 = "CHNG-COVID",
                             AR3DVCLI3 = "DV-CLI",
                             AR3FBCLI3 = "CTIS-CLIIC",
                             AR3GSSAA3_Zero = "Google-AA")) %>%
  select(-ae, -POPESTIMATE2019)

# Join sanity check
nrow(ours) == nrow(ours2)
ours = ours2

our_models <- ours %>% filter(forecaster != "Baseline")
baseline <- ours %>% filter(forecaster == "Baseline") %>%
  select(-forecaster) %>% rename(strawman_wis = wis)
our_models <- left_join(our_models, baseline)

# Join sanity check
nrow(our_models) == nrow(ours) - sum(ours$forecaster == "Baseline")

@@

>>>>

# Load our models ---------------------------------------------------------
```{r}
library(tidyverse)
ours <- readRDS(here::here("data", "results_honest_states.RDS")) %>%
  filter(forecaster != "AR3GSSAA3_Subset")
pop <- covidcast::state_census %>% select(ABBR, POPESTIMATE2019) %>%
  mutate(geo_value = tolower(ABBR)) %>% select(-ABBR)
# scale from prop to num
ours2 <- left_join(ours, pop, by = "geo_value") %>%
  mutate(wis = wis, #* POPESTIMATE2019, #/ 1e5 * 7,
         forecaster = recode(forecaster,
                             AR3 = "AR",
                             AR3CHCLI3 = "CHNG-CLI",
                             AR3CHCOV3 = "CHNG-COVID",
                             AR3DVCLI3 = "DV-CLI",
                             AR3FBCLI3 = "CTIS-CLIIC",
                             AR3GSSAA3_Zero = "Google-AA")) %>%
  select(-ae, -POPESTIMATE2019)

# Join sanity check
nrow(ours) == nrow(ours2)
ours = ours2
```

```{r}
our_models <- ours %>% filter(forecaster != "Baseline")
baseline <- ours %>% filter(forecaster == "Baseline") %>%
  select(-forecaster) %>% rename(strawman_wis = wis)
our_models <- left_join(our_models, baseline)

# Sanity check: left join
nrow(our_models) == nrow(ours) - sum(ours$forecaster == "Baseline")

# Sanity check: missingness
our_models %>%
  group_by(forecaster) %>%
  summarize(sum(is.na(wis)))

our_models %>% summarize(sum(is.na(strawman_wis)))

@@

```{r}
p1 <- ours %>% 
  group_by(forecaster, ahead) %>%
  summarize(mean_wis = Mean(wis)) %>%
  ggplot(aes(x = ahead, y = mean_wis, color = forecaster)) + 
  geom_line() + geom_point() +   
  xlab("Days ahead") + ylab("Mean WIS") + 
  scale_color_manual(values = fcast_colors, guide = guide_legend(nrow = 1)) +
  theme_bw() +
  theme(legend.position = "bottom", legend.title = element_blank())

p2 <- ours %>% 
  group_by(forecaster, ahead) %>%
  summarize(geomean_wis = GeoMean(wis)) %>%
  ggplot(aes(x = ahead, y = geomean_wis, color = forecaster)) + 
  geom_line() + geom_point() +   
  xlab("Days ahead") + ylab("Geo mean WIS") + 
  scale_color_manual(values = fcast_colors, guide = guide_legend(nrow = 1)) +
  theme_bw() +
  theme(legend.position = "bottom", legend.title = element_blank())

### 2021-10-29 forecast/paper/paper.tex
First push done
- Rewrite sig statement
- Rewrite abstract
- Pass over and rewrite muhc of intro
- Generally trying to describe the sience as much as possible,
  while maintaining language that I believe is accurate and not
  over-stating our conclusions

% Please add a significance statement to explain the relevance of your work
\significancestatement{Forecasting is a critical element in the public health
  response to any fast-moving epidemic or pandemic.  Predicting the future
  spread of some disease process, based on disease activity in the recent
  past---say, using autoregressive modeling from the time series literature---is 
  a basic and widely-used method.  In this paper we study whether including
  auxiliary indicators as additional features in such an autoregressive model
  can be useful in COVID-19 case forecasting and hotspot prediction.  We find
  that indicators based on de-identified medical insurance claims, self-reported
  symptoms through online surveys, and COVID-related internet search activity
  all provide a nontrivial improvement in forecast accuracy, though their
  utility depends on the pandemic's local dynamics (upswing, downswing, or flat
  period) at prediction time. 
}

@@

\begin{abstract}
  Reliable, short-term forecasts of traditional public health reporting streams
  (such as cases, hospitalizations, and deaths)
  ought to be a key
  % are a key
  ingredient in
  effective public health decision-making during a pandemic. Since April 2020,
  our research group has worked with data partners to collect, curate, and make
  publicly available numerous real-time COVID-19 indicators, providing multiple
  views of pandemic activity.  This paper studies the utility of these
  indicators from a forecasting perspective. We focus on five indicators,
  derived from medical insurance claims data, web search queries, and online
  survey responses. For each indicator, we ask whether its inclusion in a simple
  model leads to improved predictive accuracy relative to a similar model
  excluding it. We consider both probabilistic forecasting of confirmed COVID-19 
  case rates and binary prediction of case ``hotspots''.  Since the values of 
  indicators (and case rates) are commonly revised over time, we take special
  care to ensure that the data provided to a forecaster is the version that
  would have been available at the time the forecast was made.   
  % We demonstrate how less careful backtesting can lead to misleading
  % retrospective evaluations.   
  Our  analysis shows that consistent but modest gains in predictive accuracy 
  are obtained by using these indicators, and furthermore, these gains are
  related to periods in which the auxiliary indicators behave as ``leading
  indicators'' of case rates.     
\end{abstract}

@@

\dropcap{T}racking and forecasting indicators from public health reporting
streams---such as confirmed cases and deaths in the COVID-19 pandemic---is
crucial for understanding disease spread, correctly formulating public policy responses,
and rationally planning future public health resource needs.  A companion paper
\cite{Reinhart:2021} describes our research group's (Delphi's) efforts in
curating and maintaining a database of real-time indicators that track COVID-19
activity and other relevant phenomena. The signals (a term we use synonomously
with ``indicators'') in this database are accessible through the COVIDcast API
\cite{CovidcastAPI}, with associated R \cite{CovidcastR} and Python
\cite{CovidcastPy} packages for convenient data fetching and processing
tools. In the current paper, we quantify the utility provided by a core
set of these indicators for two fundamental prediction tasks: probabilistic
forecasting of COVID-19 case rates and prediction of future COVID-19 case
hotspots (defined by the event that a relative increase in COVID-19 cases
exceeds a certain threshold).

At the outset, we should be clear that our intent in this paper is \textit{not}
to provide an authoritative take on cutting-edge COVID-19 forecasting methods.
Instead, we provide a rigorous, quantitative assessment of
the utility that several auxiliary indicators---such as those derived from
internet surveys or medical insurance claims---provide in tasks that involve
predicting future trends in confirmed COVID-19 cases. To assess such utility in
as simple terms as possible, we center our study in the framework of a basic
autoregressive model (in which COVID cases in the near future are predicted from  
a linear combination of COVID cases in the near past), and ask whether the 
inclusion of an auxiliary indicator as an additional feature in such a model
improves its predictions.

While forecasting carries a rich literature that offers a wide range of
techniques, see, e.g., \cite{Hyndman:2018},
we purposely constrain ourselves to 
very simple models, avoiding common enhancements such as order selection,
correction of outliers/anomalies in the data, and inclusion of regularization or
nonlinearities. For the same reason, we do not account for other factors that
may well aid in forecasting, such as age-specific effects, holiday adjustments,
public health mandates, and so on.
% Paragraph from Rob
Work in \cite{ioannidis2020a} points out many additional difficulties when it
comes to
forecasting COVID-19 case rates and deaths. These include bad data, wrong
assumptions, lack of incorporation of epidemiological features, etc.  We agree
with most of these 
and our group has found COVID-19 forecasting to be extremely challenging.
However the point of our  paper is to demonstrate the utility of  auxiliary
indicators for improving COVID-19 forecasting, and to suggest that these
indicators might be useful for future pandemics. We are not suggesting that
these indicators make COVID-19 forecasting  easy. 
Despite these obstacles, analyses of forecasts submitted
%
to the COVID-19 Forecast Hub \cite{ForecastHub} by a large community of modelers 
have shown that simple, robust models have consistently been among the
best-performing over the pandemic, including time series models similar to the ones
we use in this paper. See, e.g., \cite{Cramer:2021}, for a detailed study
of death forecasting. 
% added
Similarly, the simple autoregressive models used here for case forecasting
are competitive with top models submitted to the Forecast Hub over the same
period of time (see the Supplement for this analysis).
% Our internal evaluations suggest that the same is true (at
% least, if not to a greater extent) for case forecasting.

In the companion paper \cite{Reinhart:2021}, we analyze correlations between
various indicators and COVID case rates. These correlations are natural
summaries of the contemporaneous association between an indicator and COVID
cases, but they fall short of delivering a satisfactory answer to the question
that motivates the current article: is the information contained in an indicator
demonstrably useful for the prediction tasks we care about? For such a question,
lagged correlations (e.g., measuring the correlation between an indicator and
COVID case rates several days in the future) move us in the right direction, but
still do not move us all the way there. The question about \textit{utility for
  prediction} is focused on a much higher standard than simply asking about 
correlations; to be useful in forecast or hotspot models, an indicator must
provide relevant information that is not otherwise contained in past values of
the case rate series itself. We will assess this in the most direct way
possible: by inspecting the difference in predictive performance of simple
autoregressive models trained with and without access to past values of a
particular indicator.

There is another, more subtle issue in evaluating predictive utility that 
deserves explicit mention, as it will play a key role in our analysis.
Signals computed from surveillance streams will often be subject to  
latency and/or revision. For example, a signal based on aggregated medical
insurance claims may be available after just a few days, but it can then be
substantially revised over the next several weeks as additional claims are
submitted and/or processed late. Correlations between such a signal and case
rates calculated ``after the fact'' (i.e., computed retrospectively, using the
finalized values of this signal) will not deliver an honest answer to the    
question of whether this signal would have been useful in real time. Instead,
we build predictive models using only the data that would have been available
\textit{as of} the prediction date, and compare the ensuing predictions in terms 
of accuracy. To do so, we leverage Delphi's \texttt{evalcast} R package 
\cite{EvalcastR}, which plugs into the COVIDcast API's data versioning system,  
and facilitates honest backtesting.
%
The necessity of real-time data for honest forecast evaluations has a long history in econometrics,
\cite[see,
e.g.,][]{faust2009a,diebold1991a,patterson1995a,croushore2011a,croushore2006a,harvey1983a,mankiw1986a,trivellato1986a,clark2009a,koenig2003a,howrey1978a}
but it is far less acknowledged in public health despite its demonstrable importance.

Finally, it is worth noting that examining the importance of additional features
for prediction is a core question in inferential statistics and econometrics,
with work dating back to at least \cite{Granger:1969}. Still today, drawing
rigorous inference based on predictions, without (or with lean) assumptions, is
an active field of research from both the applied and theoretical angles;
see, e.g., \cite{Diebold:2002, McCraken:2007, Diebold:2015, Stokes:2017,
  Lei:2018, Rinaldo:2019, Williamsom:2020, Zhang:2020, Dai:2021, Fryer:2021}.
Our take on this problem is in line with much of this literature; however, in
order to avoid making any explicit assumptions, we do not attempt to make formal 
significance statements, and instead, broadly examine the stability of our
conclusions with respect to numerous modes of analysis.

>>>>

% Please add a significance statement to explain the relevance of your work
\significancestatement{Validated forecasting methodology should be a key 
  element in the public health response to any fast-moving epidemic or 
  pandemic. A widely-used model for predicting the future spread of a temporal
  process, based on past data, is an autoregressive (AR) model. While basic,
  such an AR model (properly trained) can already be competitive with top models
  in operational use for COVID-19 forecasting. In this paper, we 
  exhibit five auxiliary indicators---based on de-identified medical insurance
  claims, self-reported symptoms via online surveys, and COVID-related Google
  searches---that further improve the predictive accuracy of an AR model in
  COVID-19 forecasting. The most substantial gains appear to be in quiescent  
  times; but the Google search indicator appears to also offer improvements
  during upswings in pandemic activity.   
}

@@

\begin{abstract}
  Short-term forecasts of traditional public health reporting streams (such as
  cases, hospitalizations, and deaths) are a key input to public health
  decision-making during a pandemic. Since early 2020, our research group has
  worked with data partners to collect, curate, and make publicly available
  numerous real-time COVID-19 indicators, providing multiple views of pandemic
  activity in the U.S.  This paper studies the utility of five such
  indicators---derived from de-identified medical insurance claims,
  self-reported symptoms from online surveys, and COVID-related Google search
  activity---from a forecasting perspective. For each indicator, we ask whether
  its inclusion in an autoregressive (AR) model leads to improved predictive
  accuracy relative to the same model excluding it. Such an AR model, without
  external features, is already competitive with many top COVID-19 forecasting
  models in use today. Our analysis reveals that (a) inclusion of each of these
  five indicators improves on the overall predictive accuracy of the AR model;
  (b) predictive gains are in general most pronounced during times in which
  COVID cases are trending in ``flat'' or ``down'' directions; (c) one
  indicator, based on Google searches, seems to be particularly helpful
  during ``up''  trends.      
\end{abstract}

@@

\dropcap{T}racking and forecasting indicators from public health reporting
streams---such as confirmed cases and deaths in the COVID-19 pandemic---is
crucial for understanding disease spread, correctly formulating public policy
responses, and rationally planning future public health resource needs.  A
companion paper \cite{Reinhart:2021} describes our research group's efforts,
beginning in April 2020, in curating, and maintaining a database of real-time
indicators that track COVID-19 activity and other relevant phenomena. The
signals (a term we use synonomously with ``indicators'') in this database are
accessible through the COVIDcast API \cite{CovidcastAPI}, as well as associated
R \cite{CovidcastR} and Python \cite{CovidcastPy} packages, for convenient data  
fetching and processing. In the current paper, we quantify the utility provided
by a core set of these indicators for two fundamental prediction tasks:
probabilistic forecasting of COVID-19 case rates and prediction of future
COVID-19 case hotspots (defined by the event that a relative increase in
COVID-19 cases exceeds a certain threshold).

At the outset, we should be clear that our intent in this paper is \textit{not}
to provide an authoritative take on cutting-edge COVID-19 forecasting methods.
Similarly, some authors, e.g., \cite{ioannidis2020a}, have pointed out numerous
mishaps of forecasting during the pandemic and it is not our general intent to
fix them here. Instead, we start with a basic and yet reasonably effective
predictive model for predicting future trends in COVID-19 cases, and present a
rigorous, quantitative assessment of the added value provided by auxiliary
indicators, that are derived from data sources that operate outside of 
traditional public health streams. In particular, we consider five indicators
derived from de-identified medical insurance claims, self-reported symptoms from
online surveys, and COVID-related Google searches.

To assess this value in as direct terms as possible, we base our study around a
very simple basic model: an autoregressive model, in which COVID cases in the
near future are predicted using a linear combination of COVID cases in the near
past. Forecasting carries a rich literature, offering a wide range of
sophisticated techniques, see, e.g., \cite{Hyndman:2018} for a review; however,
we purposely avoid enhancements such as order selection, correction of
outliers/anomalies in the data, and inclusion of regularization or
nonlinearities. Similarly, we do not account for other factors that may well aid
in forecasting, such as age-specific effects, holiday adjustments, and the
effects of public health mandates. All that said, despite its simplicity, the
basic autoregressive model that we consider in this paper exhibits competitive
performance (see the Supplementary Materials for details) with many of the top
COVID-19 case forecasters submitted to the U.S.\ COVID-19 Forecast Hub
\cite{ForecastHub}, which is the official source of forecasts used in public
communications by the U.S. CDC. The strong performance of the autoregressive
model here is in line with the fact that simple, robust models have also
consistently been among the best-performing ones for COVID-19 death
forecasting \cite{Cramer:2021}.

In the companion paper \cite{Reinhart:2021}, we analyze correlations between
various indicators and COVID case rates. These correlations are natural
summaries of the contemporaneous association between an indicator and COVID
cases, but they fall short of delivering a satisfactory answer to the question
that motivates the current article: is the information contained in an indicator
demonstrably useful for the prediction tasks we care about? For such a question,
lagged correlations (e.g., measuring the correlation between an indicator and
COVID case rates several days in the future) move us in the right direction, but
still do not move us all the way there. The question about \textit{utility for
  prediction} is focused on a much higher standard than simply asking about 
correlations; to be useful in forecast or hotspot models, an indicator must
provide relevant information that is not otherwise contained in past values of
the case rate series itself. We assess this by inspecting the difference in
predictive performance of simple autoregressive models trained with and without
access to past values of a particular indicator.

We find that each of the five indicators we consider--- two based on
COVID-related outpatient visits from medical insurance claims, one on
self-reported symptoms from online surveys, and one on Google searches for
anosmia or ageusia---provide an overall improvement in accuracy when
incorporated into the autorgressive model. This is true both for COVID-19 case 
forecasting and hotspot prediction. Further analysis reveals that the gains in
accuracy depend on the pandemic's dynamics at prediction time: the biggest  
gains in accuracy appear during times in which cases are ``flat'' or trending
``down''; but the indicator based on Google searches offers a most notable
improvement when cases are trending ``up''.

Careful handling of data revisions plays a key role in our analysis. Signals 
computed from surveillance streams are often subject to latency and/or 
revision. For example, a signal based on aggregated medical insurance claims may
be available after just a few days, but it can then be substantially revised
over the next several weeks as additional claims are submitted and/or processed
late. Correlations between such a signal and case rates calculated ``after the
fact'' (i.e., computed retrospectively, using the finalized values of this
signal) will not deliver an honest answer to the question of whether this signal
would have been useful in real time. Instead, we build predictive models using
only the data that would have been available \textit{as of} the prediction date,
and compare the ensuing predictions in terms of accuracy. The necessity of
real-time data for honest forecast evaluations has been recognized in
econometrics for a long time, see
\cite{howrey1978a, harvey1983a, mankiw1986a, trivellato1986a, diebold1991a,
  patterson1995a, koenig2003a, croushore2006a, clark2009a, faust2009a,
  croushore2011a}; but it appears to be far less acknowledged in public health,
despite its demonstrable importance.

Finally, it is worth noting that examining the importance of additional features
for prediction is a core question in inferential statistics and econometrics,
with work dating back to at least \cite{Granger:1969}. Still today, drawing
rigorous inference based on predictions, without (or with lean) assumptions, is
an active field of research from both the applied and theoretical angles;
see, e.g., \cite{Diebold:2002, McCraken:2007, Diebold:2015, Stokes:2017,
  Lei:2018, Rinaldo:2019, Williamsom:2020, Zhang:2020, Dai:2021, Fryer:2021}.
Our take in the current work is in line with much of this literature; however,
in order to avoid making any explicit assumptions, we do not attempt to make
formal  significance statements, and instead, broadly examine the stability of
our conclusions with respect to numerous modes of analysis.
